---
layout: default
title: Summary on CMC and Benchmark
#featimg: density-1.png
---

Which individuals have been sequenced?

For the answer see the `src/sequenced-individuals.sh` as well as the `data/dnalib/get-cmc-individuals` scripts.

```{r engine="bash"}
indivs=$HOME/projects/bsm/results/2018-09-12-sequenced-individuals/sequenced-individuals
cat $indivs
```

Thus the number of sequenced individuals is

```{r engine="bash", echo=FALSE}
indivs=$HOME/projects/bsm/results/2018-09-12-sequenced-individuals/sequenced-individuals
wc -w $indivs
```

## Extension: sequenced samples

For any given individual three types of sample are possible: *muscle*, *NeuN+* and *NeuN-*.  But not all sample types have been sequenced for every individual.  The table below---generated by `sequenced-samples.sh`---presents the details.

```{r engine="bash"}
samples=$HOME/projects/bsm/results/2018-09-12-sequenced-individuals/sequenced-samples
cat $samples
```

## Summary

```{r}
fl <- c(cmc = "~/projects/bsm/results/2018-09-12-sequenced-individuals/summary-cmc.tsv",
        cs = "~/projects/bsm/tables/summary-common-sample.tsv",
        bm = "~/projects/bsm/tables/summary-benchmark.tsv")
summ <- lapply(fl, read.delim, row.names = 1)
summ
```

Write results to file

```{r}
summ.all <- do.call(cbind, summ)
write.csv(summ.all, "~/projects/bsm/results/2018-09-12-sequenced-individuals/summary-all.csv")
```

The printed inventory on the hard disks received from the sequencing company is accessible [here](https://photos.app.goo.gl/ttjSFEHPcSV1Wrvt5).  It shows that the samples that are missing from server `ada` are also missing from the printed inventory, therefore these samples are missing because haven't been sequenced yet.  Note that the contents of one of the hard drives has been verified to match the printed inventory.
