---
layout: default
title: CEPH/Utah Illumina Call Sets
tags: [ ceph, utah, vcf, truthset, callset ]
featimg: "venn-S1-1.png"
---

Characterization of the call sets for the CEPH/Utah grandparents in Illumina BaseSpace.  The grandparents are individual NA12889, NA12890, NA12891, NA12892.  There are 3 callsets for each individual, which I refer to here as calltypes: S1, genome, and SV.  The VCF header for S1 and genome are essentially identical; both calltypes corresponds to SNPs and short indels called by Strelka.  But "genome" contains a lot more calls (VCF records) than "S1".  SV is clearly for structural variants.

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(lme4)
library(lattice)
library(latticeExtra)
opts_chunk$set(dpi = 144)
opts_chunk$set(out.width = "600px")
opts_chunk$set(dev = c("png", "pdf"))
lattice.options(default.args = list(as.table = TRUE))
lattice.options(default.theme = "standard.theme")
```

```{r message=FALSE}
library(VennDiagram)
library(VariantAnnotation)
source("2018-05-01-ceph-vars-greally.R")
source("../../src/vcf.R")
```

In this analysis only S1 and SV calltypes are studied.  The genome calltype comes with too large callsets to be manageable (memory requirement).

Import VCFs and retain only the name of each record:

```{r cache=TRUE, warning=FALSE}
indiv <- paste0("NA128", 89:92)
names(indiv) <- indiv
subdir <- c("S1", "genome", "SV")
subdir <- subdir[-2] # speed things up
names(subdir) <- subdir
vcf <- lapply(subdir, function(s) lapply(indiv, read1vcf, subdir = s))
```

Get partition sizes:

```{r cache=TRUE}
part <- do.call(rbind, lapply(subdir, function(s) data.frame(calltype = s, partition.sizes(vcf[[s]]))))
```

Set some graphic parameters for Venn diagrams:

```{r}
my.par <- list(main.cex = 1.8, fill = trellis.par.get("superpose.line")$col[seq_along(indiv)], col = "gray", cat.cex = 1.4)
```

```{r venn-S1}
grid.draw(venn.diagram(vcf[["S1"]], NULL, main = "S1", main.cex = my.par$main.cex, fill = my.par$fill, col = my.par$col, cat.cex = my.par$cat.cex))
```

```{r venn-SV}
grid.draw(venn.diagram(vcf[["SV"]], NULL, main = "SV", main.cex = my.par$main.cex, fill = my.par$fill, col = my.par$col, cat.cex = my.par$cat.cex))
```

Plot partition sizes

```{r part-sizes}
psize.plot <- xyplot(callsets.containing.partition ~ log10(calls.in.partition) | calltype, data = part, pch = "|", col = "red", grid = TRUE, cex = 2)
psize.plot
```

