---
layout: default
#title: 
featimg: density-1.png
---

Annotations in VCF files provide information for variant calling and filtering.  Therefore (joint) distribution of annotations across candidate variants (VCF records) is of interest.  Since each variant caller produces its own set of annotations, those annotations must be examined separately for each caller.

## Introduction

VariantMetaCaller finds the optimal filtering *automatically* given a set of annotations in VCFs and a chosen FDR (precision threshold).  Typically there are numerous annotations in a VCF file and even more annotations can be derived from those (see annotations generated by the `annotatevcf.sh` script using [vcflib](https://github.com/vcflib/vcflib)).  Using all possible annotations for VariantMetaCaller is not necessary because not all of them carry information for VariantMetaCaller.  Neither it is practical because of several reasons: First, the training (fitting) of support vector machines integral to VariantMetaCaller would be problematic due to correlations between annotations; second, it would require considerable reformatting of the input VCFs to comply with VariantMetaCaller; third, it would compromise the running time of VariantMetaCaller.  Following the practice established by the authors of VariantMetaCaller ([Gézsi et al](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4625715/)) our task is to *manually* select the annotations that are informative for VariantMetaCaller in the sense that those annotations are necessary for the maximum performance of VariantMetaCaller.  Therefore we need to estimate the precision-recall curve under several sets of selected annotations until we find the set of annotations both necessary and sufficient for maximum performance of VariantMetaCaller.

In summary, there are two steps:

1. variable selection (selection of annotations)
    * manual
    * configuration of VariantMetaCaller
2. training (of support vector machines)
    * automatic
    * running of VariantMetaCaller

## Preparations

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(lme4)
library(lattice)
library(VennDiagram)
opts_chunk$set(dpi = 144)
opts_chunk$set(out.width = "600px")
opts_chunk$set(dev = c("png", "pdf"))
lattice.options(default.args = list(as.table = TRUE))
lattice.options(default.theme = "standard.theme")
```

Making a list of data frames based on a list of callers.  Each data frame contains a random subsample of VCF records and a subset of VCF fields.

```{r engine="bash", eval=FALSE}
# SNVs
cd $HOME/projects/bsm/results/2018-07-03-vcf-annotations/benchmark-mix1a-mix3a/snvs
for vcf in vcf/*.vcf.gz; do vcfinfo2tsv $vcf& done
# indels
cd $HOME/projects/bsm/results/2018-07-03-vcf-annotations/benchmark-mix1a-mix3a/indels
for vcf in vcf/*.vcf.gz; do vcfinfo2tsv $vcf& done
```

```{r cache=TRUE}
callers <- c("strelka2Somatic", "strelka2Germline", "lofreqSomatic", "somaticSniper", "Tnseq")
names(callers) <- callers
# helper function
import.subsample <- function(caller, ssize = 1e4, indir = "../../results/2018-07-03-vcf-annotations/benchmark-mix1a-mix3a/snvs/") {
    df <- read.delim(paste0(indir, caller, ".vcf.gz.info.tsv"), na.strings = c("NA", "."))
    mycols <- sapply(df, function(y) ! all(is.na(y)))
    mysample <- sample(seq.int(nr <- nrow(df)), size = min(nr, ssize), replace = FALSE)
    df[mysample, mycols]
}
info <- list() # list of lists
info$snvs <- lapply(callers, import.subsample, ssize = 1e4, indir = "../../results/2018-07-03-vcf-annotations/benchmark-mix1a-mix3a/snvs/")
info$indels <- lapply(callers, import.subsample, ssize = 1e4, indir = "../../results/2018-07-03-vcf-annotations/benchmark-mix1a-mix3a/indels/")
```

```{r engine="bash", eval=TRUE}
cd $HOME/projects/bsm/results/2018-07-03-vcf-annotations/benchmark-mix1a-mix3a/
./parse-header-info
cp info-fields.csv $HOME/lab-notebook/assets/projects/bsm/
```

See [this table]({{ site.baseurl}}/assets/projects/bsm/info-fields.csv) for annotations for each caller.

Reshaping data frames into long format.

```{r}
myreshape <- function(caller, infolist = info$snvs) {
    df <- infolist[[caller]]
    df <- df[sapply(df, is.numeric)]
    foo <- function(key) data.frame(Annotation = key, Value = df[[key]])
    l <- lapply(names(df), foo)
    do.call(rbind, l)
}
info.long <- list()
info.long$snvs <- lapply(callers, myreshape, info$snvs)
info.long$indels <- lapply(callers, myreshape, info$indels)
```

## Marginal distributions

### SNVs

```{r density}
lapply(callers, function(y) densityplot(~ Value | Annotation, data = info.long$snvs[[y]], pch = ".", main = y, scales = list(relation = "free")))
```

## Pairwise joint distributions

### SNVs

```{r splom}
mysplom <- function(caller, infolist = info$snvs) {
    df <- infolist[[caller]]
    df <- df[sapply(df, is.numeric)]
    splom(df, pch = '.', main = caller)
}
lapply(callers["somaticSniper" != callers], mysplom, info$snvs)
```

