---
layout: default
title: Swapped sample labels and contamination
tags: [ strelka2, mutect2, tnseq, lofreq, somaticsniper ]
featimg: "sequence-only-2.png"
---

Swapped sample labels and contamination is tested for our sequencing data for the BSM project from CMC individuals.  This analysis was carried out first mistakenly using Individual_IDs instead of Genotyping_Sample_IDs as a basis of comparison between our BAMs and the CommonMind's previous genotyping results.  Later the mistake was discovered and the analysis was repeated with the correct Genotyping_Sample_IDs and even more BAMs.  These newer results are presented first followed by the older results from the first round of analysis.

## Introduction

Sample mislabeling (swaps) and contamination is examined using [verifyBamID](https://genome.sph.umich.edu/wiki/VerifyBamID).  Quote from the documentation:

> verifyBamID is a software that verifies whether the reads in particular file match previously known genotypes for an individual (or group of individuals), and checks whether the reads are contaminated as a mixture of two samples. verifyBamID can detect sample contamination and swaps when external genotypes are available. When external genotypes are not available, verifyBamID still robustly detects sample swaps. 

## Preparation

Synapse's [command line client](https://python-docs.synapse.org/build/html/CommandLineClient.html) was used to download BED, BIM and FAM files for preimputed CMC genotype data.  [PLINK 1.9](https://www.cog-genomics.org/plink/1.9/) was used to recode these files to VCF.  Then the VCF was compared by `verifyBamID` to each of the BAM files of our BSM project that come from CMC individuals.

```{r engine="bash", eval=FALSE}
# to be run on Ada
cd /projects/bsm/attila/results/2018-12-18-verifyBamID
synapse get syn16816490 # CMC_Human_SNP_mergedMetadata.csv
./getBED.sh # download CMC-preimputed.{bed,bim,fam}
./recode2vcf # recode CMC-preimputed.bed to CMC-preimputed.vcf
bcftools index --csi CMC-preimputed.vcf.gz # CSI index
mkdir results && cd results # subdirectory for verifyBamID results
./doCMCverifyBamID # run verifyBamID on all CMC bam files of the BSM project
# ...wait several hours until all verifyBamID processes exit...
./aggregate-selfSM
# transfer results to local workstation for further analysis with R
```

On the configuration of `verifyBamID`

1. the `--ignoreRG` flag was turned on so that no read group specific result was produced (no `XXX.selfRG` and `XXX.depthRG` files)
1. by default the `--self` flag was on so that `verifyBamID` 
   > does not try to compare the sequence reads to identify the best matching individual (which is possible with `--best` option). It only compares with the external genotypes from the same individual to the sequenced individual.---quoted from the [website](https://genome.sph.umich.edu/wiki/VerifyBamID)
1. the VCF has sample names like *0_MSSM_106* while our BAM files have sample names like *MSSM106_NeuN_pl*. This difference is taken care of by the `CMCverifyBamID` script that is eventually called by `doCMCverifyBamID`.

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(lme4)
library(lattice)
library(latticeExtra)
opts_chunk$set(dpi = 144)
opts_chunk$set(out.width = "600px")
opts_chunk$set(dev = c("png", "pdf"))
lattice.options(default.args = list(as.table = TRUE))
lattice.options(default.theme = "standard.theme")
```

### Effect of the --best option on runtime

The script below runs `verifyBamID` twice on the same input BAM: once with the `--best` option and once without it (i.e. with the `--self` option) and records runtime.

```{r engine="bash", eval=FALSE}
# to be run on Ada
/projects/bsm/attila/results/2018-12-18-verifyBamID/results.best_opt/doCMCverifyBamID.best_opt
```

The runtime is nearly four times longer with the `--best` option.

```{r engine="bash", eval=TRUE}
cd /home/attila/projects/bsm/results/2018-12-18-verifyBamID/results.best_opt
tail -n3 best.log self.log
```

Note that the input BAM is a relatively small BAM compared to others in this project so the runtime for those larger BAMs is typically longer, too.

## Results: Second round of analysis

For details see [Interpreting output files](https://genome.sph.umich.edu/wiki/VerifyBamID#Interpreting_output_files) in the `verifyBamID` website. In short, the concatenated `XXX.selfSM` files are inspected here, which describe how the reads of the sequenced sample (in the BAM file) matches the corresponding genotyped sample in the VCF

```{r}
# import concatenated selfSM files (the output file of aggregate-selfSM)
selfSM.2 <- read.delim("~/projects/bsm/results/2018-12-18-verifyBamID/results/all.selfSM")
```

### verifyBamID: Sequence-only model

The sequence-only model infers contamination based solely on the BAM file without referring to the "external" genotype in the VCF file.  Minus the log of $$f(y \mid \hat{\theta}) / f(y \mid \theta_0)$$, which is the likelihood ratio of the maximum likelihood of the unconstrained model accounting for contamination and the constrained null model of no contamination.  The larger the that value is the more likely is that the sample has been contaminated.

A strongly related quantity is *FREEMIX*, the "sequence-only estimate of contamination (0-1 scale)".  The section [A guideline to interpret output files](https://genome.sph.umich.edu/wiki/VerifyBamID#A_guideline_to_interpret_output_files) advices:

> When [CHIPMIX] $$\gg$$ 0.02 and/or [FREEMIX] $$\gg$$ 0.02, meaning 2% or more of non-reference bases are observed in reference sites, we recommend to examine the data more carefully for the possibility of contamination.

The plots below suggest that *MSSM_373_muscle* may contain contamination.

```{r sequence-only}
dotplot(sample ~ FREELK0 - FREELK1, data = selfSM.2)
dotplot(sample ~ FREEMIX, data = selfSM.2)
```

### verifyBamID: Sequence + array model

This model is an extension of the sequence-only model in that external genotype is also taken into account.  The quantities plotted below are the minus log likelihood ratio of the "contaminated" and the "uncontaminated" model as well as *CHIPMIX*.  These are analogous to the quantities of the sequence-only model but in this case there relationship between the two quantities is less clear.  For instance, *CHIPMIX* suggests, similarly to *FREEMIX*, that *MSSM_373_muscle* is a sole outlier but the likelihood ratio suggests otherwise.

```{r sequence-chip}
dotplot(sample ~ CHIPLK0 - CHIPLK1, data = selfSM.2)
dotplot(sample ~ CHIPMIX, data = selfSM.2)
```

## Results: First round of analysis

```{r}
# import concatenated selfSM files (the output file of aggregate-selfSM)
selfSM <- read.delim("~/projects/bsm/results/2018-12-18-verifyBamID/results/all.selfSM")
```

### verifyBamID: Sequence-only model

```{r sequence-only-1}
dotplot(sample ~ FREELK0 - FREELK1, data = selfSM)
dotplot(sample ~ FREEMIX, data = selfSM)
```

### verifyBamID: Sequence + array model

*CHIPMIX* is close to 1 for all samples, which indicates swapped sample label, which is very hard to imagine for all samples.  What is going on?  Did the sample names of the BAMs not match those in the VCF?

```{r sequence-chip-1}
dotplot(sample ~ CHIPLK0 - CHIPLK1, data = selfSM)
dotplot(sample ~ CHIPMIX, data = selfSM)
```

## Confirming results with nucleotide pileup

### Preparation

The results by `verifyBamID` suggest that all BAMs mismatch the corresponding external genotype in the VCF.  To confirm this I selected 19 SNPs (every 40,000th SNP in the VCF) and compared each BAM's [nucleotide pileup](https://en.wikipedia.org/wiki/Pileup_format) to the external genotype from the VCF.

First prepare some input files:

```{r engine="bash", eval=FALSE}
# to be run on Ada
cd /projects/bsm/attila/results/2018-12-18-verifyBamID
mkdir sample.40000 && cd sample.40000
# subsample every 40,000th SNP in VCF and keep only col 1-3 to produce regions file
bcftools view -H ../CMC-preimputed.vcf.gz | sed -n '2~40000p' | cut -f1-3 > CMC-preimputed.sample.40000
# keep only col 1-3 to produce regions file
cut -f1-2 CMC-preimputed.sample.40000 > CMC-preimputed.sample.40000.col12
# filter VCF for the selected SNPs
bcftools view --regions-file CMC-preimputed.sample.40000 -O v -o CMC-preimputed.sample.40000.vcf ../CMC-preimputed.vcf.gz 
```

Then run `samtools mpileup` in the `pileup4sample` script

```{r engine="bash", eval=FALSE}
# creating the pileup (to be run on Ada still)
# for all BAMs from MSSM CMC samples create pileup with samtools at the selected SNPs
# recall that PITT_064 and PITT_118 samples are missing from the VCF
./pileup4sample
# ...wait almost a day for the run to complete...
```

### Results

The columns of the pileup files are as follows (a genotype column has been added to the left w.r.t. the [standard pileup format](https://en.wikipedia.org/wiki/Pileup_format)
1. External genotype (from VCF)
1. Sequence identifier
1. Position in sequence (starting from 1)
1. Reference nucleotide at that position
1. Number of aligned reads covering that position (depth of coverage)
1. Bases at that position from aligned reads
1. Phred Quality of those bases, represented in ASCII with -33 offset (OPTIONAL)

Below are just 3 examples of the 22 pileup files under `/projects/bsm/attila/results/2018-12-18-verifyBamID/sample.40000/` on Ada.  For all three files the genotype does not match the pileup for all SNPs.

```{r engine="bash", eval=TRUE}
# run on attila-ThinkS
cd ~/projects/bsm/results/2018-12-18-verifyBamID/sample.40000
cat MSSM_179_muscle.pileup
```

```{r engine="bash", eval=TRUE}
# run on attila-ThinkS
cd ~/projects/bsm/results/2018-12-18-verifyBamID/sample.40000
cat MSSM_295_muscle.pileup
```

```{r engine="bash", eval=TRUE}
# run on attila-ThinkS
cd ~/projects/bsm/results/2018-12-18-verifyBamID/sample.40000
cat MSSM_215_NeuN_pl.pileup
```

## Checking the VCF of external genotypes

The results above suggest label swap; this has been corroborated by Chaggai using the same BAM files and the same VCF containing the external genotypes.  But besides label swap another explanation is that the VCF itself is incorrect either because

1. the PLINK files on Synapse (`bed`, `bim` and `fam`; Synapse IDs: syn2507223, syn2507168, syn2507166) are incorrect
1. the download from Synapse to Ada corrupted the files
1. the recoding using PLINK introduced error

The first explanation, incorrect files on Synapse, could be checked by someone who was involved in the CMC project; according to the Synapse file attributes it was Menachem Fromer who uploaded the files in 2014; recently, it was Gabriel Hoffman who referred us to the files as the resource that contains genotypes for CMC individuals.

The third explanation, PLINK error would mean some kind of bug in PLINK because the command in the `recode2vcf.sh` script is as simple as `plink --bfile CMC-preimputed --recode vcf bgz --out CMC-preimputed` and it ran without reported error.  Note that a recent version of PLINK was used for this operation: PLINK v1.90b6.7 64-bit (2 Dec 2018).

The second explanation, download error, was checked by comparing the VCFs that resulted from downloading the PLINK files with the Command line client *versus* the web client.  The comparison suggests that both method of downloading files leads to the same VCF.  Details below:

```{r engine="bash", eval=FALSE}
# rename files from 1st download (Command line client) from CMC-preimputed.* to CMC-preimputed-1.*
# rename files from 2nd download (Command line client) from CMC-preimputed.* to CMC-preimputed-2.*
# rename files from 3rd download (web client) from CMC-preimputed.* to CMC-preimputed-3.*
# compare 2nd download to the 1st
bcftools isec -p checkVCF/1vs2 CMC-preimputed-{1,2}.vcf.gz
bcftools view -H checkVCF/1vs2/0000.vcf
bcftools view -H checkVCF/1vs2/0001.vcf
# compare 3rd download to the 1st
bcftools isec -p checkVCF/1vs3 CMC-preimputed-{1,3}.vcf.gz
bcftools view -H checkVCF/1vs3/0000.vcf
bcftools view -H checkVCF/1vs3/0001.vcf
```

