---
layout: default
tags: [BAM, alignment, genome]
featimg: "coverage-wgs-1.png"
---

Coverage and other statistics on the aligned reads based on the CEPH mixture samples.

```{r echo=FALSE, warning=FALSE}
library(lme4)
library(lattice)
library(latticeExtra)
opts_chunk$set(dpi = 144)
opts_chunk$set(out.width = "700px")
opts_chunk$set(dev = c("png", "pdf"))
lattice.options(default.args = list(as.table = TRUE))
lattice.options(default.theme = "standard.theme")
```

The depth was obtained with `samtools depth` based on alignments such as `Mix1A.bam`.  Every $1000$th nucleotide position in the genome was retained and written in files such as `Mix1A.bam.1000.depth`.  The following expressions filter this by 10 to arrive at every $10,000$th position being retained.

```{r}
get.depth <- function(fname = "../../results/2018-03-12-ceph-alignment-stats/Mix1A.bam.1000.depth", filter.by = 10) {
    df <- read.delim(fname, header = FALSE, col.names = c("contig", "pos", "depth"))
    df$contig <- factor(df$contig)
    return(df[seq(from = 1, to = nrow(df), by = filter.by), ])
}
samples <- c("Mix1A", "Mix1B")
fnames <- paste0("../../results/2018-03-12-ceph-alignment-stats/", samples, ".bam.1000.depth")
names(fnames) <- samples
depths <- lapply(fnames, get.depth)
depths <- do.call(rbind, lapply(names(depths), function(x) cbind(data.frame(sample = x), depths[[x]])))
```

The following empirical densities are based on all contigs (chromosomes).

```{r coverage-wgs}
densityplot(~ depth, data = depths, groups = sample, subset = depth < 300, plot.points = FALSE, ylab = "density", par.settings = list(superpose.line = list(col = c("magenta", "darkgreen"))), auto.key = TRUE)
```

The depth is roughly evenly distributed along and across chromosomes, except---as expected---at pericentromeric repeat regions.

```{r depth-wgs, fig.asp = 1.5}
useOuterStrips(xyplot(depth ~ pos | sample * contig, data = depths, subset = contig %in% c(as.character(1:22), "X", "Y"), ylim = c(0, 300), pch = "."))
```
